### 2020-04-20 Report/aws_325ghz_study.tex
Work on the report. Polishing of text and figures. Added part in tentative channel options.

\subsection{Channel specifications}
%
Channels around 89, 166, 183 and 229\,GHz are assumed to be fixed, with name,
position and widths defined in Table~\ref{tab:fixed:chs}. AWS channels 31 to 36
will be implemented as a single front-end, but are here still considered as two
``bands'', denoted as 166 and 183\,GHz.

For 325\,GHz, it will be assumed that channels can be placed between 0.75 and
8.0\,GHz in terms of intermediate frequency (IF), and that up to 4 channels
inside this range can be implemented. There should be some gap between these
channels, above 100\,-\,200\,MHz, to facilitate the technical implementation.
For comparison, Table~\ref{tab:fixed:chs} includes also the position of the ICI
325\,GHz channels. As the ICI channels extend to 11\,GHz in IF, they will not
be considered for AWS but will be applied for demonstration and reference
purposes.

@@

\begin{table}[!b]
  \begin{minipage}[b]{0.5\linewidth}
  \centering  
  \begin{tabular}[c]{c|c|c}
    Channel & Frequency   & Bandwidth \\
    name    & [GHz] &  [MHz] \\
    \hline
    AWS-21  & \phantom{0}89.000 & 4000\\
    AWS-31  & 165.500 & 2800\\
    AWS-32  & 176.311 & 2000\\
    AWS-33  & 178.811 & 2000\\
    AWS-34  & 180.311 & 1000\\
    AWS-35  & 181.511 & 1000\\
    AWS-36  & 182.311 & \phantom{0}500\\
    AWS-41  & 229.000 & 2000\\
    \hline
  \end{tabular}
  \end{minipage}%
  \begin{minipage}[b]{0.5\linewidth}
  \centering  
  \begin{tabular}[c]{c|c|c}
    Channel & Frequency   & Bandwidth \\
    name    & [GHz] &  [MHz] \\
    \hline
    ICI-5  & 325.15$\pm$9.50 & 3000\\
    ICI-6  & 325.15$\pm$3.50 & 2400\\
    ICI-7  & 325.15$\pm$1.50 & 1600\\
    \hline
  \end{tabular}
  \end{minipage}  
  \caption{Fixed channel position and widths. The left table covers channels of
    AWS, with values taken from SoW. Channels AWS-21 to AWS-36 are at this
    moment mandatory for AWS, while AWS-41 is so far tentative. The right table
    lists the values of some channels of ICI \citep{eriksson:towar:20}.}
  \label{tab:fixed:chs}
\end{table}

\subsection{Noise}
%
The nominal reciever noise temperature, $T_r$ of each band is listed in
Table~\ref{tab:trec}. The noise of data for individual channels is calculated
as
\begin{equation}
  \sigma^i = c \frac{T^i_r+T^i_a}{\sqrt{\Delta\!f^i\Delta t}}
\end{equation}
where $\sigma^i$ is the noise standard deviation of channel $i$, $c$ is a
scaling factor (see below), $T^i_r$ is the reciever noise temperature of the
channel, $T^i_a$ is the antenna temperature of the channel, $\Delta\!f^i$ is the
bandwidth of channel and $\Delta t$ is the integration time.

@@

\subsubsection{Bulk profile database}
%
A database of atmospheric cases was created based on data from CloudSat and
ERA-Interim. The core idea of the approach is to use CloudSat reflectivities to
obtain as realistic vertical profiles of ice hydrometeors and rain as
possible. No external CloudSat retrievals are involved. Instead the CloudSat
reflectivites are mapped to properties for the passive simulations by assuming
a particle size distribution (PSD) and a particle habit, together denoted as
the particle model. The process involves an inversion of the radar data, but as
the same particle model is used for that inversion and to map ice water content
(IWC) and rain water content (RWC) to single scattering data when performing
the passive radiative transfer calculations, the impact of clouds on the
simulated AWS data is as realistic as possible. The thin clouds not detected by
CloudSat should not be of relevance for passive measurements frequencies
350\,GHz. Background data (temperature, humidity, 10\,m wind speed, \dots) were
taken from ERA-Interim, for the time and location of the CloudSat data. Liquid
water content (LWC) was taken from ERA-Interim, as CloudSat has no sensitivity
to the hydrometeor category.

This approach for generating atmospheric scenarios has been used by us since
\citet{rydberg:nonga:09}, and more recently in e.g.\
\citet{eriksson:towar:20,barlakas:three:20}. It is also used in
\citet{ekelund:using:20}, where the approach is described more in detail. In
the cited articles, the sensor's footprint has been considered in varying
degree, as ``beam filling'' must be considered at high impact of hydrometeors.
As clear-sky and weak cloud interference are in focus in this study, and to
save calculation time, instead CloudSat data were averaged over 10\,km to form
the atmospheric cases and standard 1D radiative transfer calculations are
applied on each case (instead of applying an independent beam approximation
inside a 2D or 3D atmosphere as done in cited works).

@@

\subsection{Radiative transfer simulations}
%
\subsubsection{ARTS}
%
All radiative transfer calculations are made by the Atmospheric Radiative
Transfer System (ARTS, \citet{eriksson:arts2:11,buehler:artst:18}), version
arts-2.3.????. Models and input used for absorption and hydrometeor properties
are described below. Ocean/water and land emissivities are taken from
TESSEM \citet{prigent2017sea} and TELSEM \citet{aires2011tool}, respectively.

The calculations were performed for each atmospheric cases. One ''clear-sky''
where all hydrometeors contents were set to zero, and one ``all-sky'' where IWC
and RWC derived from CloudSat reflectivities and LWC from ERA-Interim were
included. Both sets of calculations were were made by ARTS's interface to the
RT4 solver \citep{evans1995microwavec}. This ``scattering solver'' was used
also for clear-sky, to avoid a possible bias between clear-sky and all-sky for
insignificant hydrometeor contents. RT4 handles the first two elements of the
Stokes vector, that were converted to brightness temperatures for H- and V-
polarisation.

\subsubsection{Absorption models}
%
The absorption of nitrogen is modelled following \citet{pwr:93}. Nitrogen has a
significant impact at least around 325\,GHz at dry conditions. Also oxygen
absorption follows \citet{pwr:93}. Oxygen is in this study mainly of concern
for the 89\,GHz CloudSat simulations. LWC was treated to be
purely absorbing, based on \citet{ellison2007permittivity}.

For water vapour the present settings in RTTOV are applied. That is, largely
MPM89 \citep{liebe:89} is followed, but some parameters for the 22 and 183\,GHz
transitions are replaced \citep{saunders2018update,turner2019amsutran}. A basic
comparison to RTTOV (12.?) was made (with MHS as assumed sensor) and an
agreement around or better than 0.?\,K was found.

@@

For 183 and 325\,GHz the brightness temperatures increase when moving away from
the centre frequency of the transition. The ``kinks'' in the spectra correspond
to ozone transitions, that are both more frequent and stronger in the 325\,GHz
range than in the 183\,GHz one. See Fig.~1 of \citet{eriksson:towar:20} for how
the ozone transitions are distributed between the lower and upper 325\,GHz
sideband. The spectra of 183 and 325\,GHz deviate little close to the transition
frequency, showing that the two frequencies are of similar strength, though the
325\,GHz one is slightly weaker. The higher deviations further out in the wings
are due to different contribution of far wing and continuum absorption, that is
higher at 325\,GHz. The 229\,GHz brightness temperatures is higher than any
value around 183 and 325\,GHz, as expected 229\,GHz being a window channel.

In Figure~\ref{fig:tb:r050} the surface reflectivity is changed to 0.5, which
should roughly correspond to the highest reflectivity encountered at these
frequencies. The differences to Fig.~\ref{fig:tb:r000}  are small for 183 and
325\,GHz for the tropical and the two summer scenarios. On the other hand,
there is a clear impact in the wing part of 183\,GHz for the two winter
scenarios. The same is true for 340\,GHz, but the influence of the surface
reflectivity is considerably lower. [* Comment somewhere on the implications of
this for retrievals at high latitudes *] Beside for the tropical case, for
229\,GHz there is a pronounced dependency of the surface reflectivity, e.g.\
about 35\,K for mid-latitude winter.

@@

\subsection{...}
%
\dots Fig.~\ref{fig:wfuns:325ul}

\begin{figure}[!tb]
  \centering
  \includegraphics[height=65mm]{fascod_wf_325l_tro}\hspace{5mm}% 
  \includegraphics[clip,trim=43 0 0 0,height=65mm]{fascod_wf_325u_tro}
 \caption{...}
  \label{fig:wfuns:325ul}
\end{figure}

>>>>

\subsection{Channel specifications}
%
The specifications of channels around 89, 166, 183 and 229\,GHz are assumed to
be fixed, with name, position and widths defined in Table~\ref{tab:fixed:chs}.
AWS channels 31 to 36 will be implemented as a single front-end, but are here
still considered as two ``bands'', denoted as 166 and 183\,GHz, to more easily
separate the channels in the text.

For 325\,GHz, it will be assumed that channels can be placed between 0.75 and
8.0\,GHz in terms of intermediate frequency (IF), and that up to 4 channels
inside this range can be implemented. There should be some gap between these
channels, above 100\,-\,200\,MHz, to facilitate the technical implementation.
For comparison, Table~\ref{tab:fixed:chs} includes also the position of 
325\,GHz channels of ICI. As the ICI channels extend to 11\,GHz in IF, they will not
be considered for AWS but will be applied for demonstration and reference
purposes.

@@

\begin{table}[!b]
  \begin{minipage}[b]{0.5\linewidth}
  \centering  
  \begin{tabular}[c]{c|c|c}
    Channel & Frequency   & Bandwidth \\
    name    & [GHz] &  [MHz] \\
    \hline
    AWS-21  & \phantom{0}89.000 & 4000\\
    AWS-31  & 165.500 & 2800\\
    AWS-32  & 176.311 & 2000\\
    AWS-33  & 178.811 & 2000\\
    AWS-34  & 180.311 & 1000\\
    AWS-35  & 181.511 & 1000\\
    AWS-36  & 182.311 & \phantom{0}500\\
    AWS-4X  & 229.000 & 2000\\
    \hline
  \end{tabular}
  \end{minipage}%
  \begin{minipage}[b]{0.5\linewidth}
  \centering  
  \begin{tabular}[c]{c|c|c}
    Channel & Frequency   & Bandwidth \\
    name    & [GHz] &  [MHz] \\
    \hline
    ICI-5  & 325.15$\pm$9.50 & 3000\\
    ICI-6  & 325.15$\pm$3.50 & 2400\\
    ICI-7  & 325.15$\pm$1.50 & 1600\\
    \hline
  \end{tabular}
  \end{minipage}  
  \caption{Fixed channel position and widths. The left table covers channels of
    AWS, with values taken from SoW. Channels AWS-21 to AWS-36 are at this
    moment mandatory for AWS, while AWS-4X is so far tentative. The right table
    lists the values of some channels of ICI \citep{eriksson:towar:20}.}
  \label{tab:fixed:chs}
\end{table}

\subsection{Noise}
%
The nominal reciever noise temperature, $T_r$ of each band is listed in
Table~\ref{tab:trec}. The noise of data for individual channels is calculated
as
\begin{equation}
  \sigma^i = c \frac{T^i_r+T^i_a}{\sqrt{\Delta\!f^i\Delta t}}
\end{equation}
where $\sigma^i$ is the noise standard deviation of channel $i$, $c$ is a
scaling factor (see below), $T^i_r$ is the reciever noise temperature of the
channel, $T^i_a$ is the antenna temperature of the channel, $\Delta\!f^i$ is
the bandwidth (in terms of IF) of channel and $\Delta t$ is the integration
time.

@@

\subsubsection{Bulk profile database}
%
A database of atmospheric cases was created based on data from CloudSat and
ERA-Interim. The core idea of the approach is to use CloudSat reflectivities to
obtain as realistic vertical profiles of ice hydrometeors and rain as
possible. No external CloudSat retrievals are involved. Instead the CloudSat
reflectivites are mapped to properties for the passive simulations by assuming
a particle size distribution (PSD) and a particle habit, together denoted as
the particle model. The process involves an inversion of the radar data, but as,
the same particle model is used for that inversion and to map ice water content
(IWC) and rain water content (RWC) to single scattering data when performing
the passive radiative transfer calculations, the impact of clouds on the
simulated AWS data is as realistic as possible. The thin clouds not detected by
CloudSat should not be of relevance for passive measurements frequencies
350\,GHz. Background data (temperature, humidity, 10\,m wind speed, \dots) were
taken from ERA-Interim, for the time and location of the CloudSat data. Liquid
water content (LWC) was taken from ERA-Interim, as CloudSat has no sensitivity
to the hydrometeor category.

This approach for generating atmospheric scenarios has been used by us since
\citet{rydberg:nonga:09}, and more recently in e.g.\
\citet{eriksson:towar:20} and \citet{barlakas:three:20}. It is also used in
\citet{ekelund:using:20}, where the approach is described more in detail. In
the cited articles, the sensor's footprint has been considered in varying
degree, as ``beam filling'' must be considered at high impact of hydrometeors.
As clear-sky and weak cloud interference are in focus in this study, and to
save calculation time, instead CloudSat data were averaged over 10\,km to form
the atmospheric cases and standard 1D radiative transfer calculations are
applied on each case (instead of applying an independent beam approximation
inside a 2D or 3D atmosphere as done in cited works).

@@

\subsection{Radiative transfer simulations}
%
\subsubsection{ARTS}
%
All radiative transfer calculations are made by the Atmospheric Radiative
Transfer System (ARTS, \citet{eriksson:arts2:11,buehler:artst:18}), version
arts-2.3.????. Models and input used for absorption and hydrometeor properties
are described below. Ocean/water and land emissivities are taken from
TESSEM \citep{prigent2017sea} and TELSEM \citep{aires2011tool}, respectively.

Two calculations were performed for each atmospheric cases. One ''clear-sky''
where all hydrometeors contents were set to zero, and one ``all-sky'' where IWC
and RWC derived from CloudSat reflectivities and LWC from ERA-Interim were
included. Both sets of calculations were were made by ARTS's interface to the
RT4 solver \citep{evans1995microwavec}. This ``scattering solver'' was used
also for clear-sky, to avoid a possible bias between clear-sky and all-sky for
insignificant hydrometeor contents. RT4 handles the first two elements of the
Stokes vector, that were converted to brightness temperatures for H- and V-
polarisation.

\subsubsection{Absorption models}
%
The absorption of nitrogen is modelled following \citet{pwr:93}. Nitrogen has a
significant impact at least around 325\,GHz at dry conditions. Also oxygen
absorption follows \citet{pwr:93}. Oxygen is in this study mainly of concern
for the 94\,GHz CloudSat simulations. LWC was treated to be
purely absorbing, based on \citet{ellison2007permittivity}.

For water vapour the present settings in RTTOV are applied. That is, largely
MPM89 \citep{liebe:89} is followed, but some parameters for the 22 and 183\,GHz
transitions are replaced \citep{saunders2018update,turner2019amsutran}. A basic
comparison to RTTOV (12.?) was made (with MHS as assumed sensor) and an
agreement around or better than 0.?\,K was found [* expand *].

@@

For 183 and 325\,GHz the brightness temperatures increase when moving away from
the centre frequency of the transition. The ``kinks'' in the spectra correspond
to ozone transitions, that are both more frequent and stronger in the 325\,GHz
range than in the 183\,GHz one. See Fig.~1 of \citet{eriksson:towar:20} for how
the ozone transitions are distributed between the lower and upper 325\,GHz
sideband. The spectra of 183 and 325\,GHz deviate little close to the transition
frequency, showing that the two frequencies are of similar strength, though the
325\,GHz one is slightly weaker. The higher deviations further out in the wings
are due to different contribution of far wing and continuum absorption, that is
higher at 325\,GHz. The 229\,GHz brightness temperatures are higher than any
value around 183 and 325\,GHz, as expected 229\,GHz being a window channel.

In Figure~\ref{fig:tb:r050} the surface reflectivity is changed to 0.5, which
should roughly correspond to the highest reflectivity encountered at these
frequencies. The differences to Fig.~\ref{fig:tb:r000}  are small for 183 and
325\,GHz for the tropical and the two summer scenarios. On the other hand,
there is a clear impact in the wing part of 183\,GHz for the two winter
scenarios. The same is true for 325\,GHz, but the influence of the surface
reflectivity is considerably lower. [* Comment somewhere on the implications of
this for retrievals at high latitudes *] Beside for the tropical case, for
229\,GHz there is a pronounced dependency of the surface reflectivity, e.g.\
about 35\,K for mid-latitude winter.

@@

\subsection{Impact of upper end of IF band}
%
Figure~\ref{fig:wfuns:325ul} assumes that the channel placed at the high end of
the IF band is 3\,GHz wide (as for ICI, Table~\ref{tab:fixed:chs}), and shows
the RH weighting function if the outer channel ends at 11, 10, 9 or 8\,GHz. The
weighting functions for each sideband are here kept separated.

\begin{figure}[!tb]
  \centering
  \includegraphics[height=65mm]{fascod_wf_325l_tro}\hspace{5mm}% 
  \includegraphics[clip,trim=43 0 0 0,height=65mm]{fascod_wf_325u_tro}
  \caption{RH weighting functions for the lower (left) and upper (right) side
    of the 325.15\,GHz transition. All weighting functions are for a 3\,GHz
    bandwidth. The stated frequencies are the centre frequency of each. In
    terms of IF range, the channels cover 8-11, 7-10, 6-9 and 5-8\,GHz (from
    top-to-bottom in the order given in the legends). }
  \label{fig:wfuns:325ul}
\end{figure}

As expected, when the channel is moved away the centre frequency, the weighting
function moves downwards. However, moving the channel from 5\,-\,8\,GHz to
8\,-\,11\,GHz (in IF) results only a small shift, about 400 and 200\,m for the
lower and upper side, respectively. In terms of the resulting double sideband
channel the the shift would be 300\,m. Accordingly, there is little gain
inforcing the IF range to go above 8\,GHz. As a rule of thumb, a higher ratio
between highest and lowest IF is more costly to implement. Albeit small, a
negative consequence of a 8\,-\,11\,GHz channel, compared to 5\,-\,8\,GHz, is
some assymetry in the altitude response between lower and upper sideband. This
assymetry is caused by higher continuum and far wing absorption on the upper
side.

In summary, the benefit of extending the IF band above 8\,GHz is small and
placing channels above 8\,GHz seems unnecessary considering the limited budget
and development time of AWS.

\subsection{A three channel option}
%
The ICI channels give a rough RH weighting functions matching with AWS-32,
AWS-33 and AWS-34 (Fig.~\ref{fig:wf:mls:025}), and to find a tentative three
channel option for the AWS 325\,GHz channels specifications providing an even
better matching were tried out. The resulting specifications are found in
Table~\ref{tab:chs:prel} and example weighting functions are displayed in
Fig.~\ref {fig:3ch:prel}. This option has a IF range of 1.0 to 8.0\,GHz, and
leaves a 100\,MHz space between AWS-42 and AWS-43, and 200\,MHz space between
AWS-41 and AWS-42.

\begin{table}[!b]
  \begin{minipage}[b]{0.5\linewidth}
  \centering  
  \begin{tabular}[c]{c|c|c}
    Channel & Frequency   & Bandwidth \\
    name    & [GHz] &  [MHz] \\
    \hline
    AWS-41  & 325.15$\pm$6.50 & 3000\\
    AWS-42  & 325.15$\pm$3.55 & 2500\\
    AWS-43  & 325.15$\pm$1.60 & 1200\\
    \hline
  \end{tabular}
  \end{minipage}%
  \begin{minipage}[b]{0.5\linewidth}
  \centering  
  \begin{tabular}[c]{c|c|c}
    Channel & Frequency   & Bandwidth \\
    name    & [GHz] &  [MHz] \\
    \hline
    AWS-41  & 325.15$\pm$6.60 & 2800\\
    AWS-42  & 325.15$\pm$4.05 & 1900\\
    AWS-43  & 325.15$\pm$2.35 & 1300\\
    AWS-44  & 325.15$\pm$1.20 & \phantom{0}800\\
    \hline
  \end{tabular}
  \end{minipage}  
  \caption{Preliminary specifications of AWS 325\,GHz channels, in the case of
    of three (left) and four (right) channels.}
  \label{tab:chs:prel}
\end{table}

\begin{figure}[!tb]
  \centering
  \includegraphics[height=64mm]{fascod_3chopt_tro}\hspace{5mm}% 
  \includegraphics[clip,trim=43 0 0 0,height=64mm]{fascod_3chopt_mlw}
  \caption{RH weighting functions of the three channel option in
    Table~\ref{tab:chs:prel}, for the Fascod tropical (left) and mid-latiude
    winter (right) scenarios.}
  \label{fig:3ch:prel}
\end{figure}
\begin{figure}[!tb]
  \centering
  \includegraphics[height=64mm]{fascod_4chopt_tro}\hspace{5mm}% 
  \includegraphics[clip,trim=43 0 0 0,height=64mm]{fascod_4chopt_mlw}
  \caption{RH weighting functions of the four channel option in
    Table~\ref{tab:chs:prel}, for the Fascod tropical (left) and mid-latiude
    winter (right) scenarios.}
  \label{fig:4ch:prel}
\end{figure}

\subsection{A four channel option}
%
The tentative four channel option in Table~\ref{tab:chs:prel} gives a broader
altitude coverage, towards higher altitudes, and offers weighting functions
roughly equally spaced in altitude (Fig.~\ref {fig:4ch:prel}). This option has
a IF range of 0.7 to 8.0\,GHz, and leaves 200\,MHz space between AWS-41 and
AWS-42 and has 100\,MHz between the other channels.
